---
title: "Generating Qualtrics Dictionaries"
---

Working with qualtrics files can be tricky because surveys can utilizes question families that contain multiple levels of responses. As a result, variables belong to variable groups. The raw qualtrics names are not very helpful, so dictionary crosswalk files have been create to facilitate data preparation workflow. 

**Example Variable Family:** 

| QUALTRICS VARIABLE NAME | NEW NAME | LABEL |
| --- | --- | --- |
| Staff#1_1_1 | Staff_Fulltime_2021 | Number of full time staff in 2021 |
| Staff#1_2_1 | Staff_Parttime_2021 | Number of part time staff in 2021 |
| Staff#1_3_1 | Staff_Boardmmbr_2021 | Number of Board Members in 2021 |
| Staff#2_1_1 | Staff_Fulltime_2022 | Number of full time staff in 2022 |
| Staff#2_2_1 | Staff_Parttime_2022 | Number of part time staff in 2022 |
| Staff#2_3_1 | Staff_Boardmmbr_2022 | Number of Board Members in 2022 |


Some utility scripts have been written to extract variable dictionary elements from qualtrics survey file exports and convert them into a basic crosswalk file. 

## Qualtrics File to Crosswalk

```{r, eval=F}
# source( "../data-dictionaries/R/00-data-processing-utils.R" )

URL <- "https://raw.githubusercontent.com/UrbanInstitute/nccs-nptrends/main/data-dictionaries/R/00-data-processing-utils.R"
source( URL )

###########
###########   DATA DICTIONARY
###########

# USE LEGACY = TRUE
# FOR THE YEAR 2 DATA DICTIONARY,
# FALSE FOR SUBSEQUENT YEARS 

d <- 
  read_survey( 
    "data-raw/wave-02-qualtrics-download-29mar23.csv", 
    legacy = T )

head( as.data.frame(d) )

# exports data dictionary 

dd <- extract_colmap( d )  

# add group variables
# and factor labels 

dd <- 
  dd %>% 
  mutate( group_name = append_groups(qname) ) %>%
  group_by( group_name ) %>% 
  mutate( group_n = n(),
          group_name = ifelse( group_n > 1, group_name, "" ),
          is_group = ifelse( group_n > 1, "1", "0" ),
          group_levels = ifelse( group_n > 1, get_categories(description), "" ) ) %>%
  ungroup() %>% 
  mutate( type = ifelse( group_n > 1, "factor", "" ) ) %>% 
  select( qname, is_group, group_name, group_n, group_levels, 
          description, main, sub )


write.csv( dd, "../data-dictionaries/dd-nptrends-wave-02.csv" )
```


## Dictionary Files 

After generating the skeleton of the crosswalk file from the script, it would be completed by a research assistant by providing information for fields marked **user** in the table: 

| DD VARIABLE | DESCRIPTION | SOURCE |
| --- | --- | --- |
| q | question number (order) | qualtrix |
| vname_raw | variable name (from qualtrix export) | qualtrix |
| vname | variable name (final) | user |
| vlabel | variable label  | user |
| type | data type (numeric, character, factor, logical, date) | user |
| group | group name | r script |
| group_lev1 | factor levels | user |
| group_lev2 | second factor level for double-grouped variables (e.g. finances_1_1) | user |
| group_lev_draft | parsed categories (clean up and use as group_lev labels) | r script |
| add_noise | add noise to this variable to anonymize?  | user |
| description | survey question full | qualtrix |
| main | survey question sub | qualtrix |
| sub | survey question response categories (kindof) | qualtrix |

These steps need to be completed each year. Most of the completed dictionaries can be reused if questions are not changing, but note that changes to the order of the questions in the survey can change the qualtrics naming conventions, and any new questions added would need documentation. 

These dictionary crosswalk files are utilized in subsequent steps. 

### Preview 

```{r}
dd <- readxl::read_xlsx( "../data-dictionaries/dd-nptrends-wave-02.xlsx", sheet = "data dictionary" )

head( dd[30:50,] ) 
```

