---
title: "Summary Statistics"
execute:
  warning: false
  message: false
---

```{r, include=FALSE, echo=FALSE}
library( haven )
library( dplyr )
library( tidyr )
library( epoxy )
library( memisc )
library( labelled )

# SOURCE PREVIOUS STEPS 
knitr::purl(  "22-year2-fix-demographics.qmd"  )    # convert QMD to R script
source(       "22-year2-fix-demographics.R"    )    # run all chunks in prior step
file.remove(  "22-year2-fix-demographics.R"    )    # delete script after execution
```




  
```{r, results="asis", echo=F, eval=F}
all.questions <- c(
  program_change_qns_bool, program_change_qns_txt,
  people_served_qns_bool, people_served_qns_int,
  demand_fct_qns,
  staff_qns_bool, staff_qns_int, staff_qns_text,
  volimportance_qns_fct, donimportance_qns_fct,
  leadership_chng_qns_bool, leadership_chng_qns_text,
  race_gender_qns_bool, race_gender_qns_text,
  finance_chng_qns_bool, finance_chng_qns_text,
  cares_qns_bool, cares_qns_num,
  reserve_qns_bool, reserve_qns_num,
  finance_revenue_qns_num, finance_revenue_qns_text,
  fundraise_skrcv_qns_bool,
  fundraise_donor_qns_int,
  fundraise_change_qns_fct,
  fundraise_qns_bool,fundraise_qns_text,
  majorgift_qn_num,
  extaffairs_qns_fct,
  primary_cncrn_qn_text )

omit.these.groups <- c("id","DROP","MainAddress","Comments")
all.vars <- dd$vname[ ! dd$group %in% omit.these.groups ] |> na.omit()
show_html( codebook( survey_df[ all.vars ] ))
```






<br>
<hr>
<br>

- **`r length( c(program_change_qns_bool, program_change_qns_txt) ) `** questions about **CHANGES TO PROGRAMS AND SERVICES**
- **`r length( c(people_served_qns_bool,people_served_qns_int) ) `** questions about the **NUMBER OF PEOPLE EACH ORGANIZATION SERVES**
- **`r length( c(demand_fct_qns) ) `** question about **OVERALL PROGRAM DEMAND**
- **`r length( c(staff_qns_bool, staff_qns_int,  staff_qns_text) ) `** questions about **STAFF NUMBERS**
- **`r length( c(volimportance_qns_fct, donimportance_qns_fct) ) `** questions about **DONOR AND VOLUNTEER IMPORTANCE**
- **`r length( c(leadership_chng_qns_bool, leadership_chng_qns_text) ) `** questions about **CHANGES TO LEADERSHIP**
- **`r length( c(race_gender_qns_bool, race_gender_qns_text) ) `** questions about **THE RACE AND GENDER OF CEOS AND BOARD CHAIRS**
- **`r length( c(finance_chng_qns_bool, finance_chng_qns_text) ) `** questions about **CHANGES TO ORGANIZATIONAL FINANCES**
- **`r length( c(cares_qns_bool, cares_qns_num) ) `** questions about **CARES FUNDING**
- **`r length( c(reserve_qns_bool, reserve_qns_num) ) `** questions about **FINANCIAL RESERVES**
- **`r length( c(finance_revenue_qns_num, finance_revenue_qns_text) ) `** questions about **REVENUE SOURCES**
- **`r length( c(fundraise_skrcv_qns_bool) ) `** questions about **FUNDRAISING SOURCES**
- **`r length( c(fundraise_donor_qns_int) ) `** questions about **DONOR TYPES IN FUNDRAISING**
- **`r length( c(fundraise_change_qns_fct) ) `** questions about **FUNDRAISING YIELDS**
- **`r length( c(fundraise_qns_bool, fundraise_qns_text) ) `** questions about **FUNDRAISING STRATEGY CHANGES**
- **`r length( c(majorgift_qn_num) ) `** questions about **MAJOR GIFT AMOUNTS**
- **`r length( c(extaffairs_qns_fct) ) `** questions about **EXTERNAL AFFAIRS**  
- **`r length( c(primary_cncrn_qn_text) ) `** questions about **FUTURE CONCERNS**



<br>
<hr>
<br>


# CHANGES TO PROGRAMS AND SERVICES

```{r, results="asis", echo=F}
COLUMNS <- c(program_change_qns_bool,program_change_qns_txt)
show_html( codebook( survey_df[ COLUMNS ] ))
```


# NUMBER OF PEOPLE EACH ORGANIZATION SERVES

```{r, results="asis", echo=F}
COLUMNS <- c(people_served_qns_bool,people_served_qns_int)
show_html( codebook( survey_df[ COLUMNS ] ))
```


# OVERALL PROGRAM DEMAND

```{r, results="asis", echo=F}
COLUMNS <- demand_fct_qns
show_html( codebook( survey_df[ COLUMNS ] ))
```


# STAFF NUMBERS

```{r, results="asis", echo=F}
COLUMNS <- c(staff_qns_bool, staff_qns_int,  staff_qns_text)
show_html( codebook( survey_df[ COLUMNS ] ))
```



# DONOR AND VOLUNTEER IMPORTANCE

```{r, results="asis", echo=F}
COLUMNS <- c(volimportance_qns_fct, donimportance_qns_fct)
show_html( codebook( survey_df[ COLUMNS ] ))
```


# CHANGES TO LEADERSHIP

```{r, results="asis", echo=F}
COLUMNS <- c(leadership_chng_qns_bool, leadership_chng_qns_text)
show_html( codebook( survey_df[ COLUMNS ] ))
```



# THE RACE AND GENDER OF CEOS AND BOARD CHAIRS

```{r, results="asis", echo=F}
COLUMNS <- c(race_gender_qns_bool, race_gender_qns_text)
show_html( codebook( survey_df[ COLUMNS ] ))
```

# CHANGES TO ORGANIZATIONAL FINANCES

```{r, results="asis", echo=F}
COLUMNS <- c(finance_chng_qns_bool, finance_chng_qns_text)
show_html( codebook( survey_df[ COLUMNS ] ))
```

# CARES FUNDING

```{r, results="asis", echo=F}
COLUMNS <- c(cares_qns_bool, cares_qns_num)
show_html( codebook( survey_df[ COLUMNS ] ))
```

# FINANCIAL RESERVES

```{r, results="asis", echo=F}
COLUMNS <- c(reserve_qns_bool, reserve_qns_num)
show_html( codebook( survey_df[ COLUMNS ] ))
```

# REVENUE SOURCES

```{r, results="asis", echo=F}
COLUMNS <- c(finance_revenue_qns_num, finance_revenue_qns_text)
show_html( codebook( survey_df[ COLUMNS ] ))
```

# FUNDRAISING SOURCES**

```{r, results="asis", echo=F}
COLUMNS <- c(fundraise_skrcv_qns_bool)
show_html( codebook( survey_df[ COLUMNS ] ))
```

# DONOR TYPES IN FUNDRAISING

```{r, results="asis", echo=F}
COLUMNS <- c(fundraise_donor_qns_int)
show_html( codebook( survey_df[ COLUMNS ] ))
```

# FUNDRAISING YIELDS

```{r, results="asis", echo=F}
COLUMNS <- c(fundraise_change_qns_fct)
show_html( codebook( survey_df[ COLUMNS ] ))
```


# FUNDRAISING STRATEGY CHANGES

```{r, results="asis", echo=F}
COLUMNS <- c(fundraise_qns_bool, fundraise_qns_text)
show_html( codebook( survey_df[ COLUMNS ] ))
```

# MAJOR GIFT AMOUNTS

```{r, results="asis", echo=F}
COLUMNS <- c(majorgift_qn_num)
show_html( codebook( survey_df[ COLUMNS ] ))
```

# EXTERNAL AFFAIRS

```{r, results="asis", echo=F}
COLUMNS <- c(extaffairs_qns_fct)
show_html( codebook( survey_df[ COLUMNS ] ))
```

# FUTURE CONCERNS

```{r, results="asis", echo=F}
COLUMNS <- c(primary_cncrn_qn_text)
show_html( codebook( survey_df[ COLUMNS ] ))
```







## Save outputs


```{r}
fpath <- "DATA-PREP/02-year-two/03-data-final/"
fname <- "YEAR-02-DATA-RESTRICTED-V02.csv"
write.csv( survey_df, paste0( fpath, fname ), row.names=F, na="" )
```


```{r}
fpath <- "DATA-PREP/02-year-two/02-data-intermediate/"
fname <- "DATA-CODEBOOK.txt"
memisc::Write( codebook(survey_df), file=paste0( fpath, fname ) )
```







<style>

h1, h2 {
  margin-top: 80px;
}

.codebook-entry {
   background-color: beige;
   margin-top: 60px;
}

</style>