---
title: "Preparing Public Use Files Y2"
execute:
  warning: false
  message: false
---



## Load Restricted Data

```{r}
library( dplyr )

fpath  <- "DATA-PREP/02-year-two/03-data-final/"
fname  <- "YEAR-02-DATA-RESTRICTED-V02.csv"
df     <- read.csv( paste0( fpath, fname ) )
```



## Remove Identifying Variables

Remove all variables that: 

* identify the organization 
* identify individuals 
* include long-form responses that might include details the compromise privacy 

```{r}

drop.these <- 
  c("EIN", "EIN2", "Name_Primary", "Name_Secondary", 
    "Addr_StreetLocation", "Addr_CityState", "Addr_ZIP", 
    "ContactInfoUpdate_1", "ContactInfoUpdate_2", 
    "ContactInfoUpdate_3", "Comments", 
    "FirstName_Y2_Updated", "LastName_Y2_Updated", 
    "Title_Y2_Updated", "PrimaryConcern",
    "PrgSrvc_OthChng_Text","Staff_Other_Text_2022",
    "Staff_Other_Text_2021",
    "Staff_Other_Est_NA", "FndRaise_Text", 
    "Finance_Rev_Oth_Text","FinanceChng_Text",
    "LeadershipChng_Oth_Text", 
    "CEOrace_Text","BChairrace_Text" )

keep <- 
  c("ResponseID","year2wt","PrgSrvc_IncrNum", 
    "PrgSrvc_DcrsNum", "PrgSrvc_Suspend", "PrgSrvc_IncrSrvc", "PrgSrvc_DcrsSrvc", 
    "PrgSrvc_NewOffc", "PrgSrvc_ClsdOffc", "PrgSrvc_IncrArea", "PrgSrvc_DcrsArea", 
    "PrgSrvc_IncrFee", "PrgSrvc_DcrsFee", "PrgSrvc_ShiftOnline", 
    "PrgSrvc_AddOnline", "PrgSrvc_Oth", "PplSrv_NumServed", 
    "PplSrv_NumWait", "PplSrv_NumServed_NA_X", "PplSrv_NumWait_NA_X", 
    "Dmnd_NxtYear", "Staff_Fulltime_2021", "Staff_Parttime_2021", 
    "Staff_Boardmmbr_2021", "Staff_RegVlntr_2021", "Staff_EpsdVltnr_2021", 
    "Staff_AmerVlntr_2021", "Staff_PdCnslt_2021",  
    "Staff_Other_Est_2021", "Staff_Fulltime_2022", "Staff_Parttime_2022", 
    "Staff_Boardmmbr_2022", "Staff_RegVlntr_2022", "Staff_EpsdVltnr_2022", 
    "Staff_AmerVlntr_2022", "Staff_PdCnslt_2022",  
    "Staff_Other_Est_2022", "Staff_Fulltime_NA", "Staff_Parttime_NA", 
    "Staff_Boardmmbr_NA", "Staff_RegVlntr_NA", "Staff_EpsdVltnr_NA", 
    "Staff_AmerVlntr_NA", "Staff_PdCnslt_NA",  
    "Staff_Other_Text_NA", "VolImportance", "DonImportance", "FndRaise_DvlpVirtual", 
    "FndRaise_IncrExp", "FndRaise_DcrsExp", "FndRaise_IncStaff", 
    "FndRaise_DcrsStaff", "FndRaise_IncrCnslt", "FndRaise_DcrsCnslt", 
    "FndRaise_IncrVolntr", "FndRaise_DcrsVolntr", "FndRaise_Othr", 
    "FndRaise_MajGift_Amt", "FndRaise_Overall_Chng", 
    "FndRaise_Cashbelow250_Chng", "FndRaise_Cashabove250_Chng", "FndRaise_MajGift_Chng", 
    "FndRaise_Corp.Found_Grnt_Chng", "FndRaise_RstrGift_Chng", "FndRaise_UnrstrGift_Chng", 
    "FndRaise_DnrBlw250", "FndRaise_DnrAbv250", "FndRaise_LocGvtGrnt_Seek", 
    "FndRaise_StateGvtGrnt_Seek", "FndRaise_FedGvtGrnt_Seek", "FndRaise_LocGvtCntrct_Seek", 
    "FndRaise_StateGvtCntrct_Seek", "FndRaise_FedGvtCntrct_Seek", 
    "FndRaise_PFGrnt_Seek", "FndRaise_CFGrnt_Seek", "FndRaise_DAF_Seek", 
    "FndRaise_Corp.Found_Grnt_Seek", "FndRaise_UntdWy_Seek", "FndRaise_CombFedCmpgn_Seek", 
    "FndRaise_OthrGvngPrgrm_Seek", "FndRaise_LocGvtGrnt_Rcv", "FndRaise_StateGvtGrnt_Rcv", 
    "FndRaise_FedGvtGrnt_Rcv", "FndRaise_LocGvtCntrct_Rcv", "FndRaise_StateGvtCntrct_Rcv", 
    "FndRaise_FedGvtCntrct_Rcv", "FndRaise_PFGrnt_Rcv", "FndRaise_CFGrnt_Rcv", 
    "FndRaise_DAF_Rcv", "FndRaise_Corp.Found_Grnt_Rcv", "FndRaise_UntdWy_Rcv", 
    "FndRaise_CombFedCmpgn_Rcv", "FndRaise_OthrGvngPrgrm_Rcv", "Finance_Rev_GovtMain", 
    "Finance_Rev_Prtcpnt", "Finance_Rev_Govt3rdParty", "Finance_Rev_IndvDon", 
    "Finance_Rev_Gift", "Finance_Rev_Grnt", "Finance_Rev_Spnsr", 
    "Finance_Rev_Oth",  "Reserves_Est", "Reserves_NA_X", 
    "CARES_Rcv", "CARES_Rcv_Est", "FinanceChng_Reserves", "FinanceChng_Borrow", 
    "FinanceChng_DcrsBnft", "FinanceChng_IncrBnft", "FinanceChng_IncrExp", 
    "FinanceChng_DcrsExp", "FinanceChng_Oth",  
    "LeadershipChng_RetCEO", "LeadershipChng_RsgnCEO", "LeadershipChng_TrmnCEO", 
    "LeadershipChng_HireCEO", "LeadershipChng_IntrmCEO", "LeadershipChng_ChngBC", 
    "LeadershipChng_LostBoardMem", "LeadershipChng_RplcBoardMem", 
    "LeadershipChng_AddBoardMem", "LeadershipChng_Oth", 
    "CEOrace_AAPI", "CEOrace_Black", "CEOrace_Hisp", "CEOrace_NativeAm", 
    "CEOrace_White", "CEOrace_Oth",  "CEOgender_Man", 
    "CEOgender_Woman", "CEOgender_Trans", "CEOgender_NB", "CEOgender_Oth", 
    "CEOgender_specify", "BChairrace_AAPI", "BChairrace_Black", "BChairrace_Hisp", 
    "BChairrace_NativeAm", "BChairrace_White", "BChairrace_Oth", 
     "BChairgender_Man", "BChairgender_Woman",  
    "BChairgender_Trans", "BChairgender_NB", "BChairgender_Oth", 
    "BChairgender_specify", "ExtAffairs_GenEd", "ExtAffairs_Media", 
    "ExtAffairs_Advocacy", "ExtAffairs_GovtRs", "ExtAffairs_DiscGovtGrnt", 
    "ExtAffairs_InfoReq", "ExtAffairs_Testify", "ExtAffairs_Lobby", 
    "ExtAffairs_OrgPrtst", "ExtAffairs_Mobilize", "ExtAffairs_Petition", 
    "ExtAffairs_VoteReg", "ExtAffairs_VoteEd", 
    "CEOrace", "CEOrace_Bi", "BChairrace", "BChairrace_Bi",
    "CEOgender", "BChairgender" )

df <- df[keep]
```


## Add Noise to Numeric Responses  

Add noise to numeric responses that could possibly be compared with public records to identify the organization. 

```{r}

financials <- 
  c( "Staff_Fulltime_2021", "Staff_Parttime_2021", "Staff_Boardmmbr_2021", 
     "Staff_RegVlntr_2021", "Staff_EpsdVltnr_2021", "Staff_AmerVlntr_2021", 
     "Staff_PdCnslt_2021",  "Staff_Other_Est_2021", 
     "Staff_Fulltime_2022", "Staff_Parttime_2022", "Staff_Boardmmbr_2022", 
     "Staff_RegVlntr_2022", "Staff_EpsdVltnr_2022", "Staff_AmerVlntr_2022", 
     "Staff_PdCnslt_2022",  "Staff_Other_Est_2022",
     "FndRaise_MajGift_Amt", "FndRaise_DnrBlw250", "FndRaise_DnrAbv250",
     "Finance_Rev_GovtMain", "Finance_Rev_Prtcpnt", 
     "Finance_Rev_Govt3rdParty", "Finance_Rev_IndvDon", "Finance_Rev_Gift", 
     "Finance_Rev_Grnt", "Finance_Rev_Spnsr", "Finance_Rev_Oth",  
     "Reserves_Est", "CARES_Rcv_Est" ) 

add_noise <- function(x){ 
  if( is.na(x) | is.nan(x) ){ return(NA) }
  v <- x/20 # standard dev of 5% of base
  x <- rnorm( n=1, mean=x, sd=v ) |> round(0)
  return(x)
}

add_noise( 3000 )

make_noise <- function(x){
  x <- purrr::map_int( x, add_noise )
  return(x)
}

df2 <- df # make a copy for comparison 

df[financials] <- lapply( df[financials], make_noise ) 

max <- quantile( df$Finance_Rev_IndvDon, 0.9, na.rm=T )

plot( df$Finance_Rev_IndvDon, df2$Finance_Rev_IndvDon, 
      bty="n", pch=19, col=gray(0.5,0.5), cex=1.5,
      xlim=c(0,max), ylim=c(0,max), 
      main="Inspect Noise" )

max <- quantile( df$Staff_Fulltime_2021, 0.9, na.rm=T )

plot( df$Staff_Fulltime_2021, df2$Staff_Fulltime_2021, 
      bty="n", pch=19, col=gray(0.5,0.5), cex=1.5, 
      xlim=c(0,max), ylim=c(0,max),
      main="Inspect Noise" )
```

## Save to File 


```{r}
fpath  <- "DATA-PREP/02-year-two/03-data-final/"
fname  <- "YEAR-02-DATA-PUF.csv"
write.csv( df, paste0( fpath, fname ), row.names=FALSE, na="" )
```

