---
title: "Preparing Public Use Files Y3"
execute:
  warning: false
  message: false
---



## Load Restricted Data

```{r}
library( dplyr )

fpath  <- "DATA-PREP/03-year-three/03-data-final/"
fname  <- "YEAR-03-DATA-RESTRICTED-V04.csv"
df     <- read.csv( paste0( fpath, fname ) )
```



## Remove Identifying Variables

Remove all variables that: 

* identify the organization 
* identify individuals 
* include long-form responses that might include details the compromise privacy 

```{r}
drop.these <- 
  c("ExternalReference", "ResponseID", "RecipientLastName", "RecipientFirstName", 
    "RecipientEmail", "Completion_Status", "FiscalYear", "FiscalYear_Other", 
    "Name_Primary", "Name_Secondary", "Q122", "Q122_3_TEXT", "Addr_StreetLocation", 
    "Addr_CityState", "Addr_ZIP",
    "PrgSrvc_OthChng_Text", "Staff_Other_Text_2022", "Staff_Other_Text_2023", 
    "Staff_Other_Est_NA", "FndRaise_Text", "Finance_Rev_Oth_Text", 
    "FinanceChng_Text", "LeadershipChng_Oth_Text", "CEOrace_Text", 
    "BChairrace_Text", "Regulations_Other_TEXT", "CEOrace_Text_v2022", 
    "BChairrace_Text_v2022",
    "ContactUpdateName", "ContactUpdateTitle", 
    "ContactUpdateEmail", "Comments")
    
keep <- 
  c("PrgSrvc_IncrNum", "PrgSrvc_DcrsNum", "PrgSrvc_Suspend", "PrgSrvc_IncrSrvc", 
    "PrgSrvc_DcrsSrvc", "PrgSrvc_NewOffc", "PrgSrvc_ClsdOffc", "PrgSrvc_IncrArea", 
    "PrgSrvc_DcrsArea", "PrgSrvc_IncrFee", "PrgSrvc_DcrsFee", "PrgSrvc_ShiftOnline", 
    "PrgSrvc_AddOnline", "PrgSrvc_Oth", "PplSrv_NumServed", "PplSrv_NumWait", 
    "PplSrv_NumServed_NA_X", "PplSrv_NumWait_NA_X", "Dmnd_NxtYear", 
    "Staff_Fulltime_2022", "Staff_Parttime_2022", "Staff_Boardmmbr_2022", 
    "Staff_RegVlntr_2022", "Staff_EpsdVltnr_2022", "Staff_AmerVlntr_2022", 
    "Staff_PdCnslt_2022", "Staff_Other_Est_2022", "Staff_Fulltime_2023", 
    "Staff_Parttime_2023", "Staff_Boardmmbr_2023", "Staff_RegVlntr_2023", 
    "Staff_EpsdVltnr_2023", "Staff_AmerVlntr_2023", "Staff_PdCnslt_2023", 
    "Staff_Other_Est_2023", "Staff_Fulltime_NA", "Staff_Parttime_NA", 
    "Staff_Boardmmbr_NA", "Staff_RegVlntr_NA", "Staff_EpsdVltnr_NA", 
    "Staff_AmerVlntr_NA", "Staff_PdCnslt_NA",  "Staff_Other_Text_NA",
    "VolImportance", "DonImportance", "FndRaise_DvlpVirtual", "FndRaise_IncrExp", 
    "FndRaise_DcrsExp", "FndRaise_IncStaff", "FndRaise_DcrsStaff", 
    "FndRaise_IncrCnslt", "FndRaise_DcrsCnslt", "FndRaise_IncrVolntr", 
    "FndRaise_DcrsVolntr", "FndRaise_Othr", "FndRaise_MajGift_Amt", 
    "FndRaise_Overall_Chng", "FndRaise_Cashbelow250_Chng", "FndRaise_Cashabove250_Chng", 
    "FndRaise_MajGift_Chng", "FndRaise_Corp_Found_Grnt_Chng", "FndRaise_RstrGift_Chng", 
    "FndRaise_UnrstrGift_Chng", "FndRaise_DnrBlw250", "FndRaise_DnrAbv250", 
    "FndRaise_LocGvtGrnt_Seek", "FndRaise_StateGvtGrnt_Seek", "FndRaise_FedGvtGrnt_Seek", 
    "FndRaise_LocGvtCntrct_Seek", "FndRaise_StateGvtCntrct_Seek", 
    "FndRaise_FedGvtCntrct_Seek", "FndRaise_PFGrnt_Seek", "FndRaise_CFGrnt_Seek", 
    "FndRaise_DAF_Seek", "FndRaise_Corp_Found_Grnt_Seek", "FndRaise_UntdWy_Seek", 
    "FndRaise_CombFedCmpgn_Seek", "FndRaise_OthrGvngPrgrm_Seek", 
    "FndRaise_LocGvtGrnt_Rcv", "FndRaise_StateGvtGrnt_Rcv", "FndRaise_FedGvtGrnt_Rcv", 
    "FndRaise_LocGvtCntrct_Rcv", "FndRaise_StateGvtCntrct_Rcv", "FndRaise_FedGvtCntrct_Rcv", 
    "FndRaise_PFGrnt_Rcv", "FndRaise_CFGrnt_Rcv", "FndRaise_DAF_Rcv", 
    "FndRaise_Corp_Found_Grnt_Rcv", "FndRaise_UntdWy_Rcv", "FndRaise_CombFedCmpgn_Rcv", 
    "FndRaise_OthrGvngPrgrm_Rcv", "Finance_Rev_GovtMain", "Finance_Rev_Prtcpnt", 
    "Finance_Rev_IndvDon", "Finance_Rev_Gift", "Finance_Rev_Grnt", 
    "Finance_Rev_Spnsr", "Finance_Rev_Oth", "Reserves_Est", "Reserves_NA_X", 
    "CARES_Rcv", "CARES_Rcv_Est", "FinanceChng_Reserves", "FinanceChng_Borrow", 
    "FinanceChng_DcrsBnft", "FinanceChng_IncrBnft", "FinanceChng_IncrExp", 
    "FinanceChng_DcrsExp", "FinanceChng_Oth", "LeadershipChng_RetCEO", 
    "LeadershipChng_RsgnCEO", "LeadershipChng_TrmnCEO", "LeadershipChng_HireCEO", 
    "LeadershipChng_IntrmCEO", "LeadershipChng_ChngBC", "LeadershipChng_LostBoardMem", 
    "LeadershipChng_RplcBoardMem", "LeadershipChng_AddBoardMem", 
    "LeadershipChng_Oth", "CEOrace_AAPI", "CEOrace_Black", "CEOrace_Hisp", 
    "CEOrace_NativeAm", "CEOrace_White", "CEOrace_Oth", "CEOgender_Man", 
    "CEOgender_Woman", "CEOgender_Trans", "CEOgender_NB", "CEOgender_Oth", 
    "CEOgender_specify", "BChairrace_AAPI", "BChairrace_Black", "BChairrace_Hisp", 
    "BChairrace_NativeAm", "BChairrace_White", "BChairrace_Oth", 
    "BChairgender_Man", "BChairgender_Woman", "BChairgender_Trans", 
    "BChairgender_NB", "BChairgender_Oth", "BChairgender_specify", 
    "ExtAffairs_GenEd", "ExtAffairs_Media", "ExtAffairs_Advocacy", 
    "ExtAffairs_GovtRs", "ExtAffairs_DiscGovtGrnt", "ExtAffairs_InfoReq", 
    "ExtAffairs_Testify", "ExtAffairs_Lobby", "ExtAffairs_OrgPrtst", 
    "ExtAffairs_Mobilize", "ExtAffairs_Petition", "ExtAffairs_VoteReg", 
    "ExtAffairs_VoteEd", "Regulations", "Regulations_Local", "Regulations_State", 
    "Regulations_Federal", "Regulations_Info", "Regulations_Programming", 
    "Regulations_Forms", "Regulations_Requests", "Regulations_Advocate", 
    "Regulations_Research", "Regulations_Other", "PrimaryConcern", 
    "StaffingPlans_HireDifferent", "StaffingPlans_HireSame", "StaffingPlans_Layoff", 
    "StaffingPlans_SlowDown", "StaffingPlans_SpeedUp", "StaffingPlans_LoseVoluntary", 
    "StaffingPlans_LoseRetirement", "StaffingPlans_Retrain", "StaffingPlans_NoChange", 
    "StaffingPlans_DK", "year3wt", "Staff_Fulltime_2023_imp", 
    "Staff_Parttime_2023_imp", "Staff_Boardmmbr_2023_imp", "Staff_RegVlntr_2023_imp", 
    "Staff_EpsdVltnr_2023_imp", "Staff_AmerVlntr_2023_imp", "Staff_PdCnslt_2023_imp", 
    "CEOrace_AAPI_v2022", "CEOrace_Black_v2022", "CEOrace_Hisp_v2022", 
    "CEOrace_NativeAm_v2022", "CEOrace_White_v2022", "CEOrace_Oth_v2022", 
    "CEOgender_Man_v2022", "CEOgender_Woman_v2022", "CEOgender_Trans_v2022", 
    "CEOgender_NB_v2022", "CEOgender_Oth_v2022", "CEOgender_specify_v2022", 
    "BChairrace_AAPI_v2022", "BChairrace_Black_v2022", "BChairrace_Hisp_v2022", 
    "BChairrace_NativeAm_v2022", "BChairrace_White_v2022", "BChairrace_Oth_v2022", 
    "BChairgender_Man_v2022", "BChairgender_Woman_v2022", "BChairgender_Trans_v2022", 
    "BChairgender_NB_v2022", "BChairgender_Oth_v2022", "BChairgender_specify_v2022", 
    "CEOrace_Bi", "BChairrace_Bi", "CEOrace", "BChairrace", "CEOgender", 
    "BChairgender")

df <- df[keep]
```


## Add Noise to Numeric Responses  

Add noise to numeric responses that could possibly be compared with public records to identify the organization. 

```{r}

# identify numeric variables easily: 
# names(df)[ (lapply(df,class) |> unlist()) != "character" ] |> dput()

financials <-
   c("PplSrv_NumServed", "PplSrv_NumWait", "Staff_Fulltime_2022", 
    "Staff_Parttime_2022", "Staff_Boardmmbr_2022", "Staff_RegVlntr_2022", 
    "Staff_EpsdVltnr_2022", "Staff_AmerVlntr_2022", "Staff_PdCnslt_2022", 
    "Staff_Other_Est_2022", "Staff_Fulltime_2023", "Staff_Parttime_2023", 
    "Staff_Boardmmbr_2023", "Staff_RegVlntr_2023", "Staff_EpsdVltnr_2023", 
    "Staff_AmerVlntr_2023", "Staff_PdCnslt_2023", "Staff_Other_Est_2023", 
    "FndRaise_MajGift_Amt", "FndRaise_DnrBlw250", "FndRaise_DnrAbv250", 
    "Finance_Rev_GovtMain", "Finance_Rev_Prtcpnt", "Finance_Rev_IndvDon", 
    "Finance_Rev_Gift", "Finance_Rev_Grnt", "Finance_Rev_Spnsr", 
    "Finance_Rev_Oth", "Reserves_Est", "CARES_Rcv_Est",  
    "Staff_Fulltime_2023_imp", "Staff_Parttime_2023_imp", "Staff_Boardmmbr_2023_imp", 
    "Staff_RegVlntr_2023_imp", "Staff_EpsdVltnr_2023_imp", "Staff_AmerVlntr_2023_imp", 
    "Staff_PdCnslt_2023_imp", "CEOgender_specify_v2022")


add_noise <- function(x){ 
  if( is.na(x) | is.nan(x) ){ return(NA) }
  v <- x/20  # standard dev of 5% of base
  x <- rnorm( n=1, mean=x, sd=v ) |> round(0)
  return(x)
}

add_noise( 3000 )

make_noise <- function(x){
  x <- purrr::map_dbl( x, add_noise )
  return(x)
}

df2 <- df  # make a copy for comparison

df[financials] <- lapply( df[financials], make_noise ) 

max <- quantile( df$Finance_Rev_IndvDon, 0.9, na.rm=T )

plot( df$Finance_Rev_IndvDon, df2$Finance_Rev_IndvDon, 
      bty="n", pch=19, col=gray(0.5,0.5), cex=1.5,
      xlim=c(0,max), ylim=c(0,max), 
      main="Inspect Noise" )

max <- quantile( df$Staff_Fulltime_2022, 0.9, na.rm=T )

plot( df$Staff_Fulltime_2022, df2$Staff_Fulltime_2022, 
      bty="n", pch=19, col=gray(0.5,0.5), cex=1.5, 
      xlim=c(0,max), ylim=c(0,max),
      main="Inspect Noise" )

max <- quantile( df$Staff_Fulltime_2023, 0.9, na.rm=T )

plot( df$Staff_Fulltime_2023, df$Staff_Fulltime_2023_imp, 
      bty="n", pch=19, col=gray(0.5,0.5), cex=1.5, 
      xlim=c(0,max), ylim=c(0,max),
      main="Inspect Noise" )
```

## Save to File 


```{r}
fpath  <- "DATA-PREP/03-year-three/03-data-final/"
fname  <- "YEAR-03-DATA-PUF.csv"
write.csv( df, paste0( fpath, fname ), row.names=FALSE, na="" )
```
