---
title: "Anonymization"
---

In this section we'll work through variables with potentially identifiable information and either exclude/anonymize them.

```{r, include=FALSE, echo=FALSE}
library(dplyr)
library(readr)
library(sdcMicro)

setwd("Y:/CNP/Generosity Commission/")
survey_recode_df <- readr::read_csv("DATA-PREP/02-data-intermediate/02-wave-two/wave-02-data-intermediate-recoded.csv")
anon_df <- survey_recode_df %>% dplyr::select("EIN", "Reserves_Est")
```

# Identifying Variables for Anonymization

## Continuous Variables

Variables with numeric or integer inputs can be identifiable using publicly available 990 data. They are presented in the table below.

|Variable|Description|
|:---:|:---:|
|Reserves_Est|Dollar Value of Financial Reserves|
|CARES_Rcv_Est|Dollar Value of CARES Funding|
|Finance_Rev_GovtMain|Dollar Value of Revenue from Government Agencies|
|Finance_Rev_Prtcpnt|Dollar Value of Revenue from Self-Paying Participants|
|Finance_Rev_Govt3rdParty|Dollar Value of Revenue from Government as Third-Party Payer|
|Finance_Rev_IndvDon|Dollar Value of Revenue from Individual Donations|
|Finance_Rev_Gift|Dollar Value of Revenue from In-Kind Gifts|
|Finance_Rev_Grnt|Dollar Value of Revenue from Private Foundations/Corporate Grants|
|Finance_Rev_Spnsr|Dollar Value of Revenue from Corporate Sponsorships|
|Finance_Rev_Oth||Dollar Value of Revenue from Other Sources|

## Categorical Variables

Categorical variables on their own may not be identifiable. However, they can be combined with other variables to identify survey respondents. This possibility of composite identification is called the ["Mosaic Effect"](https://www.govinfo.gov/content/pkg/CHRG-111shrg56893/pdf/CHRG-111shrg56893.pdf).

For example, the gender/racial identities of CEOs can be used in conjunction with geographical data to identify respondents who belong to minority groups. Thus, there is a need to anonymize categorical variables that contain demographic information linked to specific survey respondents and their organizations.

A table with these variables is provided below:

|Variable(s)|Description|
|:---:|:---:|
|CEOrace_{Race} Variables|Race of Chief Executive|
|CEOgender_{Gender} Variables|Gender of Chief Executive|
|BChairrace_{Race} Variables|Race of Board Chair|
|BChairgender_{Gender} Variables|Gender of Board Chair|
|State|State for Respondent|

# Anonymizing Variables with SDCMicro

To anonymize these variables, we use the sdcMicro R package, developed by HSN, PARIS21 (OECD), Statistics Austria, The Vienna University of Technology, and the World Bank for Statistical Disclosure Control (SDC).

## Selecting Variables

sdcMicro creates an sdMicro object based on the variables needed for anyonmization process. These variables are supplied as function arguments. They are presented in the table below.

|Argument|Description|
|:---:|:---:|
|keyVars|Categorical Variables for Anonymization|
|CEOgender_{Gender} Variables|Gender of Chief Executive|
|BChairrace_{Race} Variables|Race of Board Chair|
|BChairgender_{Gender} Variables|Gender of Board Chair|
|State|State for Respondent|



```{r}
# Categorical Variables
```


## Creating a SDCMicro Object

```{r}

```

