---
title: "Anonymization"
---

In this section we'll work through variables with potentially identifiable information and either exclude/anonymize them.

```{r, include=FALSE, echo=FALSE}
library(dplyr)
library(readr)
library(sdcMicro)

setwd("Y:/CNP/Generosity Commission/")
survey_census_df <- readr::read_csv(
  "DATA-PREP/02-data-intermediate/02-wave-two/wave-02-data-intermediate-census.csv"
)
```

# Identifying Variables for Anonymization

## Continuous Variables

### Survey Variables

Variables with numeric or integer inputs can be identifiable using publicly available 990 data. They are presented in the table below.

|Variable|Description|
|:---:|:---:|
|Reserves_Est|Dollar Value of Financial Reserves|
|CARES_Rcv_Est|Dollar Value of CARES Funding|
|Finance_Rev_GovtMain|Dollar Value of Revenue from Government Agencies|
|Finance_Rev_Prtcpnt|Dollar Value of Revenue from Self-Paying Participants|
|Finance_Rev_Govt3rdParty|Dollar Value of Revenue from Government as Third-Party Payer|
|Finance_Rev_IndvDon|Dollar Value of Revenue from Individual Donations|
|Finance_Rev_Gift|Dollar Value of Revenue from In-Kind Gifts|
|Finance_Rev_Grnt|Dollar Value of Revenue from Private Foundations/Corporate Grants|
|Finance_Rev_Spnsr|Dollar Value of Revenue from Corporate Sponsorships|
|Finance_Rev_Oth||Dollar Value of Revenue from Other Sources|

We anonymize by randomly sampling and adding a small amount of noise to the data.

```{r}
cols_to_anonymize <- c(
  "CARES_Rcv_Est",
  "Finance_Rev_GovtMain",
  "Finance_Rev_Prtcpnt",
  "Finance_Rev_Govt3rdParty",
  "Finance_Rev_IndvDon",
  "Finance_Rev_Gift",
  "Finance_Rev_Grnt",
  "Finance_Rev_Spnsr",
  "Finance_Rev_Oth"
)

# Create anonymization function
anon_continuous <- function(var){
  var <- var * (1 + runif(n=1, min=-0.05, max=0.05))
  var <- var * (1 + rnorm(n=1, sd=0.05))
  return(var)
}

# Anonymize Continuous Variables
survey_anon_df <- survey_census_df %>% 
  dplyr::mutate(
    dplyr::across(
      tidyselect::all_of(cols_to_anonymize),
      ~ anon_continuous(.x)
    )
  )
```

### Census Variables

Similarly, census metadata at tract and CBSA levels can be used to identify the specific location of a nonprofit, leading to identification. We apply the same anonymization process to these columns.

```{r}
# Retrieve Census variables for anonymization
census_cols_to_anonymize <-
  c(
    "total_population_County",
    "housing_units_County",
    "occupied_County",
    "vacant_County",
    "renter_occ_County",
    "white_perc_County",
    "black_perc_County",
    "asian_perc_County",
    "hawaiian_perc_County",
    "american_alaskan_perc_County",
    "two_or_more_perc_County",
    "other_perc_County",
    "rural_perc_County",
    "bachelors_perc_County",
    "hispanic_perc_County",
    "poverty_perc_County",
    "unemployment_County",
    "turnover_perc_County",
    "med_family_income_adj_County",
    "med_gross_rent_adj_County",
    "med_household_income_adj_County",
    "median_value_adj_County",
    "total_population_Tract",
    "housing_units_Tract",
    "occupied_Tract",
    "vacant_Tract",
    "renter_occ_Tract",
    "white_perc_Tract",
    "black_perc_Tract",
    "asian_perc_Tract",
    "hawaiian_perc_Tract",
    "american_alaskan_perc_Tract",
    "two_or_more_perc_Tract",
    "other_perc_Tract",
    "rural_perc_Tract",
    "bachelors_perc_Tract",
    "hispanic_perc_Tract",
    "poverty_perc_Tract",
    "unemployment_Tract",
    "turnover_perc_Tract",
    "med_family_income_adj_Tract",
    "med_gross_rent_adj_Tract",
    "med_household_income_adj_Tract",
    "median_value_adj_Tract",
    "total_population_CBSA",
    "housing_units_CBSA",
    "occupied_CBSA",
    "vacant_CBSA",
    "renter_occ_CBSA",
    "unemployment_CBSA",
    "med_family_income_adj_CBSA",
    "med_gross_rent_adj_CBSA",
    "med_household_income_adj_CBSA",
    "median_value_adj_CBSA"
  )


# Run anonymization
# Anonymize Continuous Variables
survey_anon_df <- survey_anon_df %>%
  dplyr::mutate(
    dplyr::across(
      tidyselect::all_of(census_cols_to_anonymize),
      ~ anon_continuous(.x)
    )
  )
```


## Selecting Columns for Release in Final Data

With anonymization complete, we can remove columns with identifiable information and information not relevant to the survey before release.

```{r}
release_cols <-
  c(
    "PrgSrvc_IncrNum",
    "PrgSrvc_DcrsNum",
    "PrgSrvc_Suspend",
    "PrgSrvc_IncrSrvc",
    "PrgSrvc_DcrsSrvc",
    "PrgSrvc_NewOffc",
    "PrgSrvc_ClsdOffc",
    "PrgSrvc_IncrArea",
    "PrgSrvc_DcrsArea",
    "PrgSrvc_IncrFee",
    "PrgSrvc_DcrsFee",
    "PrgSrvc_ShiftOnline",
    "PrgSrvc_AddOnline",
    "PrgSrvc_Oth",
    "PrgSrvc_OthChng_Text",
    "PplSrv_NumServed",
    "PplSrv_NumWait",
    "PplSrv_NumServed_NA_X",
    "PplSrv_NumWait_NA_X",
    "Dmnd_NxtYear",
    "Staff_Fulltime_2021",
    "Staff_Parttime_2021",
    "Staff_Boardmmbr_2021",
    "Staff_RegVlntr_2021",
    "Staff_EpsdVltnr_2021",
    "Staff_AmerVlntr_2021",
    "Staff_PdCnslt_2021",
    "Staff_Other_Text_2021",
    "Staff_Other_Est_2021",
    "Staff_Fulltime_2022",
    "Staff_Parttime_2022",
    "Staff_Boardmmbr_2022",
    "Staff_RegVlntr_2022",
    "Staff_EpsdVltnr_2022",
    "Staff_AmerVlntr_2022",
    "Staff_PdCnslt_2022",
    "Staff_Other_Text_2022",
    "Staff_Other_Est_2022",
    "Staff_Fulltime_NA",
    "Staff_Parttime_NA",
    "Staff_Boardmmbr_NA",
    "Staff_RegVlntr_NA",
    "Staff_EpsdVltnr_NA",
    "Staff_AmerVlntr_NA",
    "Staff_PdCnslt_NA",
    "Staff_Other_Text_NA",
    "Staff_Other_Est_NA",
    "VolImportance",
    "DonImportance",
    "FndRaise_DvlpVirtual",
    "FndRaise_IncrExp",
    "FndRaise_DcrsExp",
    "FndRaise_IncStaff",
    "FndRaise_DcrsStaff",
    "FndRaise_IncrCnslt",
    "FndRaise_DcrsCnslt",
    "FndRaise_IncrVolntr",
    "FndRaise_DcrsVolntr",
    "FndRaise_Othr",
    "FndRaise_Text",
    "FndRaise_MajGift_Amt",
    "FndRaise_Overall_Chng",
    "FndRaise_Cashbelow250_Chng",
    "FndRaise_Cashabove250_Chng",
    "FndRaise_MajGift_Chng",
    "FndRaise_Corp-Found_Grnt_Chng",
    "FndRaise_RstrGift_Chng",
    "FndRaise_UnrstrGift_Chng",
    "FndRaise_DnrBlw250",
    "FndRaise_DnrAbv250",
    "FndRaise_LocGvtGrnt_Seek",
    "FndRaise_StateGvtGrnt_Seek",
    "FndRaise_FedGvtGrnt_Seek",
    "FndRaise_LocGvtCntrct_Seek",
    "FndRaise_StateGvtCntrct_Seek",
    "FndRaise_FedGvtCntrct_Seek",
    "FndRaise_PFGrnt_Seek",
    "FndRaise_CFGrnt_Seek",
    "FndRaise_DAF_Seek",
    "FndRaise_Corp-Found_Grnt_Seek",
    "FndRaise_UntdWy_Seek",
    "FndRaise_CombFedCmpgn_Seek",
    "FndRaise_OthrGvngPrgrm_Seek",
    "FndRaise_LocGvtGrnt_Rcv",
    "FndRaise_StateGvtGrnt_Rcv",
    "FndRaise_FedGvtGrnt_Rcv",
    "FndRaise_LocGvtCntrct_Rcv",
    "FndRaise_StateGvtCntrct_Rcv",
    "FndRaise_FedGvtCntrct_Rcv",
    "FndRaise_PFGrnt_Rcv",
    "FndRaise_CFGrnt_Rcv",
    "FndRaise_DAF_Rcv",
    "FndRaise_Corp-Found_Grnt_Rcv",
    "FndRaise_UntdWy_Rcv",
    "FndRaise_CombFedCmpgn_Rcv",
    "FndRaise_OthrGvngPrgrm_Rcv",
    "Finance_Rev_GovtMain",
    "Finance_Rev_Prtcpnt",
    "Finance_Rev_Govt3rdParty",
    "Finance_Rev_IndvDon",
    "Finance_Rev_Gift",
    "Finance_Rev_Grnt",
    "Finance_Rev_Spnsr",
    "Finance_Rev_Oth",
    "Finance_Rev_Oth_Text",
    "Time_FirstClick_Finances",
    "Time_LastClick_Finances",
    "Time_PageSubmit_Finances",
    "Time_ClickCount_Finances",
    "Reserves_Est",
    "Reserves_NA_X",
    "CARES_Rcv",
    "CARES_Rcv_Est",
    "FinanceChng_Reserves",
    "FinanceChng_Borrow",
    "FinanceChng_DcrsBnft",
    "FinanceChng_IncrBnft",
    "FinanceChng_IncrExp",
    "FinanceChng_DcrsExp",
    "FinanceChng_Oth",
    "FinanceChng_Text",
    "LeadershipChng_RetCEO",
    "LeadershipChng_RsgnCEO",
    "LeadershipChng_TrmnCEO",
    "LeadershipChng_HireCEO",
    "LeadershipChng_IntrmCEO",
    "LeadershipChng_ChngBC",
    "LeadershipChng_LostBoardMem",
    "LeadershipChng_RplcBoardMem",
    "LeadershipChng_AddBoardMem",
    "LeadershipChng_Oth",
    "LeadershipChng_Oth_Text",
    "CEOrace_AAPI",
    "CEOrace_Black",
    "CEOrace_Hisp",
    "CEOrace_NativeAm",
    "CEOrace_White",
    "CEOrace_Oth",
    "CEOrace_Text",
    "CEOgender_Man",
    "CEOgender_Woman",
    "CEOgender_Trans",
    "CEOgender_NB",
    "CEOgender_Oth",
    "CEOgender_specify",
    "BChairrace_AAPI",
    "BChairrace_Black",
    "BChairrace_Hisp",
    "BChairrace_NativeAm",
    "BChairrace_White",
    "BChairrace_Oth",
    "BChairrace_Text",
    "BChairgender_Man",
    "BChairgender_Woman",
    "BChairgender_Trans",
    "BChairgender_NB",
    "BChairgender_Oth",
    "BChairgender_specify",
    "ExtAffairs_GenEd",
    "ExtAffairs_Media",
    "ExtAffairs_Advocacy",
    "ExtAffairs_GovtRs",
    "ExtAffairs_DiscGovtGrnt",
    "ExtAffairs_InfoReq",
    "ExtAffairs_Testify",
    "ExtAffairs_Lobby",
    "ExtAffairs_OrgPrtst",
    "ExtAffairs_Mobilize",
    "ExtAffairs_Petition",
    "ExtAffairs_VoteReg",
    "ExtAffairs_VoteEd",
    "PrimaryConcern",
    "CEOrace_Bi",
    "BChairrace_Bi",
    "State",
    "StateAbbr",
    "County",
    "City",
    "total_population_County",
    "housing_units_County",
    "occupied_County",
    "vacant_County",
    "renter_occ_County",
    "white_perc_County",
    "black_perc_County",
    "asian_perc_County",
    "hawaiian_perc_County",
    "american_alaskan_perc_County",
    "two_or_more_perc_County",
    "other_perc_County",
    "rural_perc_County",
    "bachelors_perc_County",
    "hispanic_perc_County",
    "poverty_perc_County",
    "unemployment_County",
    "turnover_perc_County",
    "med_family_income_adj_County",
    "med_gross_rent_adj_County",
    "med_household_income_adj_County",
    "median_value_adj_County",
    "total_population_Tract",
    "housing_units_Tract",
    "occupied_Tract",
    "vacant_Tract",
    "renter_occ_Tract",
    "white_perc_Tract",
    "black_perc_Tract",
    "asian_perc_Tract",
    "hawaiian_perc_Tract",
    "american_alaskan_perc_Tract",
    "two_or_more_perc_Tract",
    "other_perc_Tract",
    "rural_perc_Tract",
    "bachelors_perc_Tract",
    "hispanic_perc_Tract",
    "poverty_perc_Tract",
    "unemployment_Tract",
    "turnover_perc_Tract",
    "med_family_income_adj_Tract",
    "med_gross_rent_adj_Tract",
    "med_household_income_adj_Tract",
    "median_value_adj_Tract",
    "cbsa_title_CBSA",
    "metro_CBSA",
    "total_population_CBSA",
    "housing_units_CBSA",
    "occupied_CBSA",
    "vacant_CBSA",
    "renter_occ_CBSA",
    "unemployment_CBSA",
    "med_family_income_adj_CBSA",
    "med_gross_rent_adj_CBSA",
    "med_household_income_adj_CBSA",
    "median_value_adj_CBSA"
  )

survey_release_df <- survey_anon_df %>% 
  dplyr::select(
    tidyselect::all_of(release_cols)
  )
```

# Saving Release Data Set

```{r}
# Save anonymized data
setwd("Y:/CNP/Generosity Commission/")
readr::write_csv(
  survey_anon_df,
  "DATA-PREP/02-data-intermediate/02-wave-two/wave-02-data-intermediate-anonymized.csv"
)

# Save release data
setwd("Y:/CNP/Generosity Commission/")
readr::write_csv(
  survey_release_df,
  "DATA-PREP/03-data-final/Public-use/nptrends-puf-wave-02.csv"
)
```

