---
title: "Anonymization"
---

In this section we'll work through variables with potentially identifiable information and either exclude/anonymize them.

```{r, include=FALSE, echo=FALSE}
library(dplyr)
library(readr)

setwd("Y:/CNP/Generosity Commission/")
survey_recode_df <- readr::read_csv("DATA-PREP/02-data-intermediate/02-wave-two/wave-02-data-intermediate-recoded.csv")
```

# Anonymizing Continuous Variables

Variables with numeric or integer inputs can be identifiable using publicly available 990 data. They are presented in the table below.

|Variable|Description|
|:---:|:---:|
|Reserves_Est|Dollar Value of Financial Reserves|
|CARES_Rcv_Est|Dollar Value of CARES Funding|
|Finance_Rev_GovtMain|Dollar Value of Revenue from Government Agencies|
|Finance_Rev_Prtcpnt|Dollar Value of Revenue from Self-Paying Participants|
|Finance_Rev_Govt3rdParty|Dollar Value of Revenue from Government as Third-Party Payer|
|Finance_Rev_IndvDon|Dollar Value of Revenue from Individual Donations|
|Finance_Rev_Gift|Dollar Value of Revenue from In-Kind Gifts|
|Finance_Rev_Grnt|Dollar Value of Revenue from Private Foundations/Corporate Grants|
|Finance_Rev_Spnsr|Dollar Value of Revenue from Corporate Sponsorships|
|Finance_Rev_Oth||Dollar Value of Revenue from Other Sources|

These can be anonymized by the creation of synthetic data that mimics the underlying distribution of these variables in the 

```{r}

```

# Anonymizing Discrete Variables

Our variables
